---
title: "Network Science Homework 5"
author: "Arghya Kannadaguli (ak5357)"
date: "2025-03-01"
output: html_document
---

**Assignment Description:** Select a real-world network, apply one of the community detection methods, and visualize the communities.

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE,
  warning = FALSE,
  #fig.width = 6,
  #fig.asp = .8,
  out.width = "90%",
  fig.align = "center")
```

```{r libraries, ecoh = FALSE}
library(igraph)
library(tidyverse)
library(ggplot2)
```

## **Zebra** Network Dataset

"This is a directed social network of bottlenose dolphins. The nodes are the bottlenose dolphins (genus Tursiops) of a bottlenose dolphin community living off Doubtful Sound, a fjord in New Zealand (spelled fiord in New Zealand). An edge indicates a frequent association. The dolphins were observed between 1994 and 2001."

_Data Source: [The Konnect Project](http://konect.cc/networks/moreno_zebra/)_

## **Data Import**

_Find a real-world network._

Here we are pulling in the physicians network described above.
```{r data_import}
iceland =
  readr::read_delim("data/iceland/out.iceland", skip = 1) |> 
  janitor::remove_empty(which = "cols") |> 
  janitor::clean_names() |> 
  as.matrix() |> 
  graph_from_edgelist()
```

## **Community Detection**

**Greedy Algorithm**

The documentation for the network says that it is undirected. However, igraph recognizes the iceland network as being directed, so we will pass it into the `cluster_fast_greedy()` function as an undirected graph, using the `as_undirected()` function.
```{r}
is_directed(iceland)
ic = cluster_fast_greedy(as_undirected(iceland))

```











