---
title: "Network Science Homework 5"
author: "Arghya Kannadaguli (ak5357)"
date: "2025-03-01"
output: html_document
---

**Assignment Description:** Select a real-world network, apply one of the community detection methods, and visualize the communities.

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE,
  warning = FALSE,
  #fig.width = 6,
  #fig.asp = .8,
  out.width = "90%",
  fig.align = "center")
```

```{r libraries, ecoh = FALSE}
library(igraph)
library(tidyverse)
library(ggplot2)
```

## **Iceland** Network Dataset

"This is a network of sexual contacts of male homosexuals in Iceland, collected in 1992."

_Data Source: [The Konnect Project](http://konect.cc/networks/moreno_iceland/)_

## **Data Import**

_Find a real-world network._

Here we are pulling in the physicians network described above.
```{r data_import}
iceland =
  readr::read_delim("data/iceland/out.iceland", skip = 1) |> 
  janitor::remove_empty(which = "cols") |> 
  janitor::clean_names() |> 
  as.matrix() |> 
  graph_from_edgelist()
```

## **Basic Analysis/Visualization**

There are **`r vcount(iceland)` nodes** and **`r ecount(iceland)` edges** in this network. Here is a visualization of the network using the FR layout. I tried the KK layout as well, but ultimately felt FR yielded a more visually organized result.

```{r}
plot(iceland,
     layout=layout_with_fr,
     vertex.label = NA,
     vertex.size = 5,
     vertex.color = "lightpink",
     edge.width = 1.5,
     edge.color = "maroon",
     edge.arrow.size = 0,
     main = "Fruchterman-Reingold (FR) Layout")
```

## **Community Detection**

**Greedy Algorithm**

The documentation for the network says that it is undirected. However, igraph recognizes the iceland network as being directed, so we will pass it into the `cluster_fast_greedy()` function as an undirected graph, using the `as_undirected()` function.
```{r}
is_directed(iceland)
ic = cluster_fast_greedy(as_undirected(iceland))
```

Using the greedy algorithm, how many communities are in the network?
```{r}
length(ic)
```

What are the sizes of these communities?
```{r}
sizes(ic)
```

Let's visualize the communities.
```{r}
plot(ic, iceland,
     vertex.size = 5,
     vertex.label = NA,
     edge.arrow.size = 0.3,
     #layout = layout_with_fr
     )
```

